---
title: "Data Cleaning and Analysis"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Required Packages

```{r}
library(tidyverse)
library(scales)
```

library(usmap)
library(maps)
library(socviz)

## Data Loading

### Retail Data

```{r}
read_csv("/Users/kenjinchang/github/seasonality-of-us-meat-consumption/data/st_proc_mn_wtfix_abb.csv.gz") %>%
  select(-...1) %>%
  head(5)
```

### Agricultural Subsidy Data

```{r}
read_csv("/Users/kenjinchang/github/seasonality-of-us-meat-consumption/data/county_data.csv") %>%
  head(5)
```

County Shapefile Data

```{r}
map_data("county") %>%
  head(6)
```

## Total Volume (kg) of Processed Meat Sales by State

```{r}
read_csv("/Users/kenjinchang/github/seasonality-of-us-meat-consumption/data/st_proc_mn_wtfix_abb.csv.gz") %>%
  select(-...1) %>%
  group_by(state) %>%
  summarize(total_kg=sum(units_kg,na.rm=TRUE)) %>% 
  ggplot(aes(x=total_kg,y=reorder(state,total_kg))) + 
  geom_col(fill="firebrick") + 
  xlab("") + 
  ylab("") +
  theme(axis.text.x=element_text(angle=90,vjust=0.5,hjust=1),panel.grid.minor=element_blank())
```

## Mean Volume (kg) of Processed Meat Sales by State

```{r}
read_csv("/Users/kenjinchang/github/seasonality-of-us-meat-consumption/data/st_proc_mn_wtfix_abb.csv.gz") %>%
  select(-...1) %>%
  group_by(state) %>%
  summarize(mean_kg=mean(units_kg,na.rm=TRUE)) %>% 
  ggplot(aes(x=mean_kg,y=reorder(state,mean_kg))) + 
  geom_col(fill="firebrick") + 
  xlab("") + 
  ylab("") +
  theme(axis.text.x=element_text(angle=90,vjust=0.5,hjust=1),panel.grid.minor=element_blank())
```

## Total Volume (kg) of Processed Meat Sales by Month

```{r}
read_csv("/Users/kenjinchang/github/seasonality-of-us-meat-consumption/data/st_proc_mn_wtfix_abb.csv.gz") %>%
  select(-...1) %>%
  group_by(mn) %>%
  summarize(total_kg=sum(units_kg,na.rm=TRUE)) %>% 
  ggplot(aes(x=mn,y=total_kg)) + 
  geom_col(fill="firebrick") + 
  scale_x_continuous(breaks=c(1:12)) +
  xlab("") + 
  ylab("") +
  theme(panel.grid.minor=element_blank())
```

### Mean Sales by Month

```{r}
read_csv("/Users/kenjinchang/github/seasonality-of-us-meat-consumption/data/st_proc_mn_wtfix_abb.csv.gz") %>%
  select(-...1) %>%
  group_by(mn) %>%
  summarize(mean_kg=mean(units_kg,na.rm=TRUE)) %>% 
  ggplot(aes(x=mn,y=mean_kg)) + 
  geom_col(fill="firebrick") + 
  scale_x_continuous(breaks=c(1:12)) +
  xlab("") + 
  ylab("") +
  theme(panel.grid.minor=element_blank())
```

## State-Level Agricultural Subsidy Payments 

```{r}
us_counties <- map_data("county")
us_counties %>%
  head(6)
```

```{r}
county_data <- read_csv("/Users/kenjinchang/github/seasonality-of-us-meat-consumption/data/county_data.csv") %>%
  mutate(county_no_state=gsub("(.*),.*","\\1",county)) %>%
  mutate(county_abb=str_remove_all(county_no_state, " County| Parish")) %>%
  mutate(state=tolower(state),county_abb=tolower(county_abb)) %>%
  unite(id,c("state","county_abb"),sep=", ",remove=FALSE) %>%
  select(id,state,county_abb,total_subsidies_1995_2021,perc_state_total) %>%
  mutate(perc_natl_total=perc_state_total/sum(perc_state_total)) %>%
  mutate(total_subsidies_1995_2021_bil=total_subsidies_1995_2021/1000000000)
county_data %>%
  head(6)
```

```{r}
county_map <- map_data("county") %>%
  unite(id,c("region","subregion"),sep=", ",remove=FALSE)
county_map %>%
  head(6)
```


```{r}
county_full <- left_join(county_map,county_data,by="id")
county_full %>%
  head(6)
```

### Total Subsidies Given across Counties, 1995-2021

```{r}
ggplot(county_full,aes(x=long,y=lat,fill=total_subsidies_1995_2021_bil,group=group)) + 
  geom_polygon(color="white",linewidth=0.05) +
  coord_map(projection="albers",lat0=39,lat1=45) + 
  scale_fill_distiller(palette="Greens",trans="reverse",na.value="white") +
  labs(fill="Total Agricultural Subsidies Given, 1995-2021") +
  xlab("") + 
  ylab("") +
  theme(legend.position="bottom",panel.grid=element_blank(),panel.background=element_blank(),axis.text=element_blank(),axis.ticks=element_blank())
```

### Percent of All Agricultural Subsidies Given across Counties, 1995-2021

```{r}
ggplot(county_full,aes(x=long,y=lat,fill=perc_natl_total,group=group)) + 
  geom_polygon(color="white",linewidth=0.05) +
  coord_map(projection="albers",lat0=39,lat1=45) + 
  scale_fill_distiller(palette="Greens",trans="reverse",na.value="white",labels=percent) +
  labs(fill="Proportion of Agricultural Subsidies Given, 1995-2021") +
  xlab("") + 
  ylab("") +
  theme(legend.position="bottom",panel.grid=element_blank(),panel.background=element_blank(),axis.text=element_blank(),axis.ticks=element_blank())
```

### Mean subsidies given by state

```{r}
county_data %>%
  group_by(state) %>%
  summarize(mean_subsidies_1995_2021=mean(total_subsidies_1995_2021)) %>%
  ggplot(aes(x=mean_subsidies_1995_2021,y=reorder(state,mean_subsidies_1995_2021))) + 
  geom_col(fill="seagreen")
```

Total subsidies given by state

```{r}
county_data %>%
  group_by(state) %>%
  summarize(state_subsidies_1995_2021=sum(total_subsidies_1995_2021)) %>%
  ggplot(aes(x=state_subsidies_1995_2021,y=reorder(state,state_subsidies_1995_2021))) + 
  geom_col(fill="seagreen")
```

## State-Level Processed Meat Sales (Per Capita)

### Mean Annual Sales Volume (kg) by State

```{r}
read_csv("/Users/kenjinchang/github/seasonality-of-us-meat-consumption/data/st_proc_mn_wtfix_abb.csv.gz") %>%
  select(-...1) %>%
  group_by(state) %>%
  summarize(mean_annual_kg=sum(units_kg,na.rm=TRUE)/3) %>%
  ggplot(aes(x=reorder(state,mean_annual_kg),y=mean_annual_kg)) + 
  geom_col(fill="firebrick") + 
  theme(axis.text.x=element_text(angle=90,vjust=0.5,hjust=1)) 
```

### Total Sales Volume (kg) by State by Year

```{r}
read_csv("/Users/kenjinchang/github/seasonality-of-us-meat-consumption/data/st_proc_mn_wtfix_abb.csv.gz") %>%
  select(-...1) %>%
  group_by(state,yr) %>%
  summarize(total_kg=sum(units_kg,na.rm=TRUE)) %>%
  mutate(yr=as.character(yr)) %>%
  ggplot(aes(x=total_kg,y=reorder(state,total_kg),fill=yr)) + 
  geom_col(position="dodge") + 
  scale_fill_brewer(palette="Reds")
```

### Total Sales Volume (kg) by Year

```{r}
read_csv("/Users/kenjinchang/github/seasonality-of-us-meat-consumption/data/st_proc_mn_wtfix_abb.csv.gz") %>%
  select(-...1) %>%
  group_by(yr) %>%
  summarize(total_kg=sum(units_kg,na.rm=TRUE)) %>%
  ggplot(aes(x=yr,y=total_kg)) + 
  geom_col(fill="firebrick")
```

### Mean Per Capita Sales Volume (kg) by State

```{r}
state_sales_data <- read_csv("/Users/kenjinchang/github/seasonality-of-us-meat-consumption/data/st_proc_mn_wtfix_abb.csv.gz") %>%
  select(-...1) %>%
  group_by(state) %>%
  summarize(mean_annual_kg=sum(units_kg,na.rm=TRUE)/3) 
state_sales_data %>%
  head(6)
```

Prep for left_join with state-level population data

Population data taken from [here](https://www.census.gov/data/tables/time-series/demo/popest/2010s-state-total.html), specifically the annual estimates of the resident population for the US

```{r}
state_pop_data <- read_csv("/Users/kenjinchang/github/seasonality-of-us-meat-consumption/data/state_pop_data.csv") %>%
  rowwise() %>%
  mutate(mean_pop=mean(c(pop2017,pop2018,pop2019))) %>%
  select(state,mean_pop)
state_pop_data %>%
  head(6)
```

```{r}
state_data <- left_join(state_sales_data,state_pop_data,by="state") %>%
  rowwise() %>%
  mutate(pc_annual_kg=mean_annual_kg/mean_pop) %>%
  mutate(state=tolower(state))
state_data %>%
  head(6)
```

```{r}
ggplot(state_data,aes(y=reorder(state,pc_annual_kg),x=pc_annual_kg)) + 
  geom_point(color="firebrick")
```

```{r}
state_map <- map_data("state") %>%
  rename(state=region)
state_map %>%
  head(6)
```


```{r}
state_full <- left_join(state_data,state_map,by="state")
state_full %>%
  head(6)
```

```{r}
ggplot(state_full,aes(x=long,y=lat,fill=pc_annual_kg,group=group)) + 
  geom_polygon(color="white",linewidth=0.05) +
  coord_map(projection="albers",lat0=39,lat1=45) + 
  scale_fill_distiller(palette="YlOrRd",trans="reverse",na.value="white") + 
  labs(fill="Mean Annual Per Capita Processed Meat Sales (kg), 2017-2019") +
  xlab("") + 
  ylab("") +
  theme(legend.position="bottom",panel.grid=element_blank(),panel.background=element_blank(),axis.text=element_blank(),axis.ticks=element_blank())
```

### Importing county-level water usage data

```{r}
county_water_data <- read_csv("/Users/kenjinchang/github/seasonality-of-us-meat-consumption/data/county_water_data.csv",skip=1) %>%
  select(STATE,COUNTY,FIPS,'IR-IrTot','IR-WGWFr','IR-WFrTo') %>%
  rename(state=STATE,county=COUNTY,fips=FIPS,acres='IR-IrTot',groundwater='IR-WGWFr',freshwater='IR-WFrTo') %>%
  mutate(county=str_remove_all(county," County| Parish")) %>%
  mutate(county=tolower(county)) %>%
  mutate(across(state,str_replace,'AL','alabama')) %>%
  mutate(across(state,str_replace,'AK','alaska')) %>%
  mutate(across(state,str_replace,'AZ','arizona')) %>%
  mutate(across(state,str_replace,'AR','arkansas')) %>%
  mutate(across(state,str_replace,'CA','california')) %>%
  mutate(across(state,str_replace,'CO','colorado')) %>%
  mutate(across(state,str_replace,'CT','connecticut')) %>%
  mutate(across(state,str_replace,'DE','delaware')) %>%
  mutate(across(state,str_replace,'FL','florida')) %>%
  mutate(across(state,str_replace,'GA','georgia')) %>%
  mutate(across(state,str_replace,'HI','hawaii')) %>%
  mutate(across(state,str_replace,'ID','idaho')) %>%
  mutate(across(state,str_replace,'IL','illinois')) %>%
  mutate(across(state,str_replace,'IN','indiana')) %>%
  mutate(across(state,str_replace,'IA','iowa')) %>%
  mutate(across(state,str_replace,'KS','kansas')) %>%
  mutate(across(state,str_replace,'KY','kentucky')) %>%
  mutate(across(state,str_replace,'LA','louisiana')) %>%
  mutate(across(state,str_replace,'ME','maine')) %>%
  mutate(across(state,str_replace,'MD','maryland')) %>%
  mutate(across(state,str_replace,'MA','massachusetts')) %>%
  mutate(across(state,str_replace,'MI','michigan')) %>%
  mutate(across(state,str_replace,'MN','minnesota')) %>%
  mutate(across(state,str_replace,'MS','mississippi')) %>%
  mutate(across(state,str_replace,'MO','missouri')) %>%
  mutate(across(state,str_replace,'MT','montana')) %>%
  mutate(across(state,str_replace,'NE','nebraska')) %>%
  mutate(across(state,str_replace,'NV','nevada')) %>%
  mutate(across(state,str_replace,'NH','new hampshire')) %>%
  mutate(across(state,str_replace,'NJ','new jersey')) %>%
  mutate(across(state,str_replace,'NM','new mexico')) %>%
  mutate(across(state,str_replace,'NY','new york')) %>%
  mutate(across(state,str_replace,'NC','north carolina')) %>%
  mutate(across(state,str_replace,'ND','north dakota')) %>%
  mutate(across(state,str_replace,'OH','ohio')) %>%
  mutate(across(state,str_replace,'OK','oklahoma')) %>%
  mutate(across(state,str_replace,'OR','oregon')) %>%
  mutate(across(state,str_replace,'PA','pennsylvania')) %>%
  mutate(across(state,str_replace,'RI','rhode island')) %>%
  mutate(across(state,str_replace,'SC','south carolina')) %>%
  mutate(across(state,str_replace,'SD','south dakota')) %>%
  mutate(across(state,str_replace,'TN','tennessee')) %>%
  mutate(across(state,str_replace,'TX','texas')) %>%
  mutate(across(state,str_replace,'UT','utah')) %>%
  mutate(across(state,str_replace,'VT','vermont')) %>%
  mutate(across(state,str_replace,'VA','virginia')) %>%
  mutate(across(state,str_replace,'WA','washington')) %>%
  mutate(across(state,str_replace,'WV','west virginia')) %>%
  mutate(across(state,str_replace,'WI','wisconsin')) %>%
  mutate(across(state,str_replace,'WY','wyoming')) %>%
  unite(id,c("state","county"),sep=", ",remove=FALSE)
county_water_data %>%
  head(6)
```

```{r}
county_water_full <- left_join(county_map,county_water_data,by="id")
county_water_full %>%
  head(6)
```

```{r}
ggplot(county_water_full,aes(x=long,y=lat,fill=freshwater,group=group)) + 
  geom_polygon(color="white",linewidth=0.05) +
  coord_map(projection="albers",lat0=39,lat1=45) + 
  scale_fill_distiller(palette="Blues",trans="reverse",na.value="white") +
  labs(fill="Total Freshwater Withdrawals (Mgal/d), 2015") +
  xlab("") + 
  ylab("") +
  theme(legend.position="bottom",panel.grid=element_blank(),panel.background=element_blank(),axis.text=element_blank(),axis.ticks=element_blank())
```

```{r}
ggplot(county_water_full,aes(x=long,y=lat,fill=groundwater,group=group)) + 
  geom_polygon(color="white",linewidth=0.05) +
  coord_map(projection="albers",lat0=39,lat1=45) + 
  scale_fill_distiller(palette="Blues",trans="reverse",na.value="white") +
  labs(fill="Total Groundwater Withdrawals (Mgal/d), 2015") +
  xlab("") + 
  ylab("") +
  theme(legend.position="bottom",panel.grid=element_blank(),panel.background=element_blank(),axis.text=element_blank(),axis.ticks=element_blank())
```

```{r}
ggplot(county_water_full,aes(x=long,y=lat,fill=acres,group=group)) + 
  geom_polygon(color="white",linewidth=0.05) +
  coord_map(projection="albers",lat0=39,lat1=45) + 
  scale_fill_distiller(palette="Blues",trans="reverse",na.value="white") +
  labs(fill="Total Acres Irrigated, 2015") +
  xlab("") + 
  ylab("") +
  theme(legend.position="bottom",panel.grid=element_blank(),panel.background=element_blank(),axis.text=element_blank(),axis.ticks=element_blank())
```

prepping health data for spatial join 

```{r}
county_health_data <- read_csv("/Users/kenjinchang/github/seasonality-of-us-meat-consumption/data/county_health_data.csv") %>%
  mutate(state=tolower(state),county=tolower(county)) %>%
  unite(id,c("state","county"),sep=", ",remove=FALSE) 
county_health_data %>%
  head(6)
```

```{r}
county_health_full <- left_join(county_map,county_health_data,by="id")
county_health_full %>%
  head(6)
```

```{r}
ggplot(county_health_full,aes(x=long,y=lat,fill=perc.obese,group=group)) + 
  geom_polygon(color="white",linewidth=0.05) +
  coord_map(projection="albers",lat0=39,lat1=45) + 
  scale_fill_distiller(palette="Reds",trans="reverse",na.value="white") +
  labs(fill="Percent of Obese Adults, 2021") +
  xlab("") + 
  ylab("") +
  theme(legend.position="bottom",panel.grid=element_blank(),panel.background=element_blank(),axis.text=element_blank(),axis.ticks=element_blank())
```

```{r}
ggplot(county_health_full,aes(x=long,y=lat,fill=perc.obese.quart,group=group)) + 
  geom_polygon(color="white",linewidth=0.05) +
  coord_map(projection="albers",lat0=39,lat1=45) + 
  scale_fill_distiller(palette="Reds",trans="reverse",na.value="white") +
  labs(fill="Quartiles of Obese Adults, 2021") +
  xlab("") + 
  ylab("") +
  theme(legend.position="bottom",panel.grid=element_blank(),panel.background=element_blank(),axis.text=element_blank(),axis.ticks=element_blank())
```

```{r}
ggplot(county_health_full,aes(x=long,y=lat,fill=prev.hosp.rate,group=group)) + 
  geom_polygon(color="white",linewidth=0.05) +
  coord_map(projection="albers",lat0=39,lat1=45) + 
  scale_fill_distiller(palette="Reds",trans="reverse",na.value="white") +
  labs(fill="Preventable Hospitalization Rate, 2021") +
  xlab("") + 
  ylab("") +
  theme(legend.position="bottom",panel.grid=element_blank(),panel.background=element_blank(),axis.text=element_blank(),axis.ticks=element_blank())
```

```{r}
ggplot(county_health_full,aes(x=long,y=lat,fill=yppl.rate,group=group)) + 
  geom_polygon(color="white",linewidth=0.05) +
  coord_map(projection="albers",lat0=39,lat1=45) + 
  scale_fill_distiller(palette="Reds",trans="reverse",na.value="white") +
  labs(fill="Years of Preventable Life Lost, 2021") +
  xlab("") + 
  ylab("") +
  theme(legend.position="bottom",panel.grid=element_blank(),panel.background=element_blank(),axis.text=element_blank(),axis.ticks=element_blank())
```

### Adding in USDA livestock data 

```{r}
county_livestock_data <- read_csv("/Users/kenjinchang/github/seasonality-of-us-meat-consumption/data/county_livestock_data.csv") %>%
  select(FIPSTEXT,y17_M110_valueNumeric,y17_M111_valueNumeric,y17_M113_valueNumeric,y17_M114_valueNumeric) %>%
  rename(fips=FIPSTEXT,chg_tot_cattle=y17_M110_valueNumeric,avg_num_cattle=y17_M111_valueNumeric,chg_milkcow=y17_M113_valueNumeric,chg_beefcow=y17_M114_valueNumeric)
county_livestock_data %>%
  head(6)
```


### Spatial Joining USDA livestock data with USGS water use and map data 

```{r}
county_livestock_water_full <- left_join(county_livestock_data,county_water_full, by='fips')
county_livestock_water_full %>%
  head(6)
```

```{r}
ggplot(county_livestock_water_full,aes(x=long,y=lat,fill=avg_num_cattle,group=group)) + 
  geom_polygon(color="white",linewidth=0.05) +
  coord_map(projection="albers",lat0=39,lat1=45) + 
  scale_fill_distiller(palette="Oranges",trans="reverse",na.value="white") +
  labs(fill="Average Number of Cattle and Calves per 100 Acres, 2017") +
  xlab("") + 
  ylab("") +
  theme(legend.position="bottom",panel.grid=element_blank(),panel.background=element_blank(),axis.text=element_blank(),axis.ticks=element_blank())
```

```{r}
ggplot(county_livestock_water_full,aes(x=long,y=lat,fill=chg_milkcow,group=group)) + 
  geom_polygon(color="white",linewidth=0.05) +
  coord_map(projection="albers",lat0=39,lat1=45) + 
  scale_fill_distiller(palette="Oranges",trans="reverse",na.value="white") +
  labs(fill="Change in Milk Cow Inventory, 2012-2017") +
  xlab("") + 
  ylab("") +
  theme(legend.position="bottom",panel.grid=element_blank(),panel.background=element_blank(),axis.text=element_blank(),axis.ticks=element_blank())
```


```{r}
ggplot(county_livestock_water_full,aes(x=long,y=lat,fill=chg_beefcow,group=group)) + 
  geom_polygon(color="white",linewidth=0.05) +
  coord_map(projection="albers",lat0=39,lat1=45) + 
  scale_fill_distiller(palette="Oranges",trans="reverse",na.value="white") +
  labs(fill="Change in Beef Cow Inventory, 2012-2017") +
  xlab("") + 
  ylab("") +
  theme(legend.position="bottom",panel.grid=element_blank(),panel.background=element_blank(),axis.text=element_blank(),axis.ticks=element_blank())
```

### Adding in USDA Economics Dataset

```{r}
county_economics_data <- read_csv("/Users/kenjinchang/github/seasonality-of-us-meat-consumption/data/county_economics_data.csv") %>%
  select(FIPSTEXT,y17_M044_valueNumeric) %>%
  rename(fips=FIPSTEXT,feedexp_prop=y17_M044_valueNumeric)
county_economics_data %>%
  head(6)
```

### Spatial Joining USDA econnomics data with USDA livestock and USGS water use and map data 

```{r}
county_economics_livestock_water_full <- left_join(county_economics_data,county_livestock_water_full, by='fips')
county_economics_livestock_water_full %>%
  head(6)
```

```{r}
ggplot(county_economics_livestock_water_full,aes(x=long,y=lat,fill=feedexp_prop,group=group)) + 
  geom_polygon(color="white",linewidth=0.05) +
  coord_map(projection="albers",lat0=39,lat1=45) + 
  scale_fill_distiller(palette="Purples",trans="reverse",na.value="white") +
  labs(fill="Proportion of 2017 expenses spent on livestock feed, 2017") +
  xlab("") + 
  ylab("") +
  theme(legend.position="bottom",panel.grid=element_blank(),panel.background=element_blank(),axis.text=element_blank(),axis.ticks=element_blank())
```


Probably also worth building in variable (y17_M044), which tabulates expenses for feed purchase as a percent of total farm production expenses in 2017

Next steps: need to add county health and county water data. Now looking to pivot so that one project focuses on how county health aligns with state retail patterns of red and processed meat and the other looks at how agricultural subsidy payments align with water usage for irrigation (and water scarcity), both at the county level 
  